---
title: "Lesson 6a. Presenting Input Data Analysis"
author: "Assoc. Prof. D. Phillips and N. Uhan"
date: "SA421 -- Simulation Modeling -- Fall 2017"
output:
  html_notebook:
    css: css/sa421.css
---

# What should be in an input data analysis report?

For __each set of input data__, your report should contain:

* A description of the "candidate" distributions that you will fit to the
  input data, including why you chose these distributions.
  
    - Use histograms, aspects of the data itself, and context (i.e., where is
      the data from) to support your selection.
      
    - For this course, we will only use the following distributions: uniform,
      normal, gamma, and exponential.
    
* An evaluation of the fit for each candidate distribution, including:

    - a graph comparing a histogram of the data with plots of the candidate
      distribution densities,
    
    - the AIC, BIC, and K-S statistic for each candidate distribution,
      presented in a table.

* Finally, a conclusion with your choice of distribution as well as the
  arguments for the choice.


# Reading data 

* Data comes in a variety of formats.

* Luckily, there are many R functions and packages that help us read data into R.

* We've already seen:

    - `read.table()` from base R,
    
    - `read_excel()` from the `readxl` package
    
* [Here is the documentation for base R functions that read different kinds of R data, including CSV files.](https://www.rdocumentation.org/packages/utils/versions/3.4.1/topics/read.table)


# Making things pretty

* Let's use the arrival data from Lesson 6 to illustrate:

```{r}
# Load readxl - package for reading Excel files
library(readxl)

# Read arrival times from Excel file
arrival.data <- read_excel('cafe-data.xlsx', range = 'Sheet1!A2:A88', col_names = 'arrival.time')

# Load dplyr
library(dplyr)

# Sort arrival times 
# Compute interarrival times in seconds, convert to numeric 
# Assume that simultaneous arrivals are in fact 1 second apart
# Remove NA data
arrival.data <- arrival.data %>%
  arrange(arrival.time) %>%
  mutate(interarrival.time = as.numeric(arrival.time - lag(arrival.time))) %>% 
  mutate(interarrival.time = if_else(interarrival.time <= 0, 1, interarrival.time)) %>%
  filter(!is.na(interarrival.time))

# Load fitdistrplus package
library(fitdistrplus)

# Find maximum likelihood estimator for exponential distribution
interarrival.fit.exp <- fitdist(arrival.data$interarrival.time, "exp")

# Find maximum likelihood estimator for gamma distribution
interarrival.fit.gamma <- fitdist(arrival.data$interarrival.time, "gamma")

# Calculate goodness-of-fit staistics/criteria
interarrival.gof <- gofstat(list(interarrival.fit.exp, interarrival.fit.gamma), 
                            fitnames = c('exponential', 'gamma'))
```

* The graphs output by the functions in the `fitdistplus` package -- like
  `denscomp` and `cdfcomp` do  nice job of creating publication-quality
  graphs.

* How can we present the maximum likelihood estimation results in a
  professional manner?

* __Don't__ simply print the summary of `fitdist` like this: 

```{r}
# Show maximum likelihood estimator for exponential distribution
summary(interarrival.fit.exp)
```

* We can get the estimate of the rate parameter above individually with:

```{r}
interarrival.fit.exp$estimate['rate']
```


* So, instead, just write a sentence explaining what maximum likelihood
  estimation finds. For example:
    
    > Assuming our data fits an exponential distribution, maximum likelihood 
    > estimation indicates that our data best fits such a distribution with a rate 
    > parameter `r interarrival.fit.exp$estimate['rate']`.
    
    - You can incorporate __any__ R variable into your R Markdown text using the technique above. 
    
* Now, how about the goodness-of-fit criteria and statistics?

* If we just print the results, it doesn't look very professional, and it includes information we are not going to use:

```{r}
print(interarrival.gof)
```

* Instead, we can create a table directly in R Markdown with the information we want. One way to do this is below.
  (You may want to extend the width of the code editor, or turn off line wrapping, to see how this code works.)

|                              | Exponential                             | Gamma                             |
| -----------------------------| --------------------------------------: | --------------------------------: |
| Kolmogorov-Smirnov statistic | `r interarrival.gof$ks['exponential']`  | `r interarrival.gof$ks['gamma']`  |
| AIC                          | `r interarrival.gof$aic['exponential']` | `r interarrival.gof$aic['gamma']` | 
| BIC                          | `r interarrival.gof$bic['exponential']` | `r interarrival.gof$bic['gamma']` | 

* [Here is the documentation on tables in R Markdown](http://rmarkdown.rstudio.com/authoring_pandoc_markdown.html#tables) -- the above table is called a "pipe table."